The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-01-27 src/ms.tex
Expanded a little the list of new features to highlight
Clarify WCS idea
lots of edits.

Major versions of the core package -- that is to say versions which add and/or
modify functionality -- are released approximately every six months, and are
then maintained with releases that fix issues, until the next major version
is released. However, every two years a major release is designated as a
long-term support (LTS) release which continues to be maintained for up to two
years \citep{ape2}. The motivation for LTS releases is to provide longer-term stable
versions of astropy that users requiring a high level of stability can make
use of if they do not always need the latest features (this could include
for example telescope pipelines and so on). The 5.0 release of the core package was designated as LTS,
and since it was released at the end of 2021, it will be maintained until the
end of 2023.

@@

\secauthor{Lía Corrales, David Shupe + Learn team}

\begin{itemize}
\item Current status and scope
\item Vision for the future (review what we said in v2.0 paper)
\item User forums and engagement with user base
\item Workshops
\end{itemize}

@@

>>>>

Summarize a few notable things in v5.0 as compared to the last LTS release.

@@

\secauthor{Lía Corrales, David Shupe, Kelle Cruz + Learn team}

\subsubsection{Current status and scope}

{\it Learn Astropy} is an umbrella term that acknowledges the broad educational efforts made by the Astropy Project, which are led by the Learn Team.
The efforts focus on developing online content and workshops covering astro-specific coding tasks in Python.
As described in \textsl{[Astropy-ii paper]}, There are four different types of Learn Astropy content: \textit{tutorials}, consisting of Jupyter Notebook lessons that are published in HTML format online; \textit{guides}, which are a series of lessons providing a foundational resource for performing a certain type of astronomical analyses; \textit{examples}, which are snippets of code that showcase a short task that can be performed with Astropy or an affiliated package; and \textit{documentation}, which is contained within the code base.
This categorization drives content development, infrastructure choices, and the appearance of the Learn Astropy website.
The Learn Team meets weekly to work on creating, expanding, improving these educational resources.

\begin{itemize}

\item {\bf Learn Astropy website:} The website has a new search interface powered by
Algolia which is used by major e-commerce websites. This work has been
supported in part by the Dunlap grant.

\item \textbf{Tutorials} Currently, we have \# of tutorials, spanning a wide range of astronomical topics and common tasks.

\item \textbf{Guide} {\it ccdproc} Jupyter book has been completed.

\item {\bf Workshops:} the Project has been conducting workshops at winter meetings
of the American Astronomical Society since AAS 225 in January 2015. Up to the
start of the coronavirus pandemic, these were full-day in-person workshops with
as many as 90 participants and a dozen facilitators from the project. During the
pandemic, these workshops were moved to an online format and split into
basic and advanced sessions. Additionally, beginning with the AAS 238 online
meeting, the workshops have been expanded to Summer AAS meetings. The Learn
team finds that the workshop audience is best found as AAS meetings as opposed
to more general Python meetings.

The Astropy Project provided another mode of community engagement at AAS Meetings 235 and 237 by organizing a NumFOCUS Sponsored Projects booth in the AAS Exhibit Hall. Funding for the exhibit hall was provided alternately by NumFOCUS and later by the Moore Foundation funding (\textbf{check this}). The booth hosted a series of Q\&A special sessions during AAS 235 and webinars during the virtual AAS 237 meetings, to provide the general astronomy community information and access to experts on a variety of open source astronomical tools.
\end{itemize}

\subsubsection{Learn vision for the future}

The Learn plan going forward is to continue to improve the tutorial website and facilate new content;
to index third-party tutorials; and to look for supportable opportunities to expand the reach of
Astropy workshops beyond the American Astronomical Society meetings.

@@

\section{Community Engagement}

\item {\bf User forums:} The Astropy Project has historically maintained a number of avenues for users to seek help or gain access to the developer community. This includes the Astropy mailing list, the Python in Astronomy Facebook group, and the Astropy Slack workspace. The Moore Foundation grant allows this Slack space to
be on a paid plan; additionally NumFOCUS has negotiated a special rate for
open-source projets. While the Slack and Astropy-dev mailing lists are primarily used to discuss the project direction and updates, it was noted that the use of a Facebook community could present a barrier to open source. The Astropy Project identified a need for a public, archived, searchable, and easily-to-navigate interface for users to ask for help Accordingly, we have commissioned
a Discourse site which is more open than the Facebook group and more user-oriented
than Astropy Slack. A benefit of the Moore Foundation grant is that Astropy
developers are able to invoice as independent contractors the time they
spend helping users on these forums.
